---
title: "Exploratory"
author: "Adam Kiehl"
date: "3/4/21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(alphavantager)
```
Import a list of assets in the S&P 500 and remove entries known to have incomplete data. 
```{r, warning=FALSE}
spList <- read.csv('./data/spList.csv') %>%
  select(-X) %>%
  filter(!(Symbol %in% c('BF.B', 'CARR', 'OTIS', 'VTRS', 'VNT')))
```
Initialize API key for Alpha Vantage communication. 
```{r}
api_key1 = 'CZUL4Z920E682K7N'
av_api_key(api_key1)
```
Use a for loop to extract time series data for each asset in spList. Trading volume and daily price highs are stored for entries since the beginning of 2016.
```{r}
volumes = list()
highs = list()
for (i in 1:nrow(spList)) {
  sym = as.character(spList[i,1])
  temp = av_get(symbol = sym, 'TIME_SERIES_DAILY', outputsize='full')
  volume_temp = list(temp$volume[which(temp$timestamp > '2016-01-01')])
  high_temp = list(temp$high[which(temp$timestamp > '2016-01-01')])
  volumes = rbind(volumes, volume_temp)
  highs = rbind(highs, high_temp)
  Sys.sleep(12)
}
```
Order time series data into columns with each variable representing a day. 
```{r}
sp_daily_volumes = data.frame(spList)
sp_daily_highs = data.frame(spList)
times = temp$timestamp[which(temp$timestamp > '2020-01-01')] 
for (i in 1:length(times)) {
  volume_temp = c()
  for (j in 1:500) {
    volume_temp = c(volume_temp, volumes[[j]][i])
  }
  sp_daily_volumes = cbind(sp_daily_volumes, volume_temp)
  high_temp = c()
  for (j in 1:500) {
    high_temp = c(high_temp, highs[[j]][i])
  }
  sp_daily_highs = cbind(sp_daily_highs, high_temp)
}
```
Rename columns according to dates. 
```{r}
col_names = c('Symbol', 'Name', 'Sector', as.character(times))
names(sp_daily_volumes) = col_names
names(sp_daily_highs) = col_names
```
Write both data frames to .csv for safe-keeping.
```{r}
write.csv(sp_daily_volumes, './data/sp_daily_volumes_lg.csv')
write.csv(sp_daily_highs, './data/sp_daily_highs_lg.csv')
```
