---
title: "Exploratory"
author: "Adam Kiehl"
date: "3/4/21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(alphavantager)
```
Import a list of assets in the S&P 500 and remove entries with incomplete data. 
```{r, warning=FALSE}
spList <- read.csv('./data/spList.csv') %>%
  select(-X) %>%
  filter(!(Symbol %in% c('AMD', 'LB', 'ROP', 'RCL')))
```
Initialize API key for Alpha Vantage communication. 
```{r}
api_key = 'CZUL4Z920E682K7N'
av_api_key(api_key)
```
Use a for loop to extract time series data for each asset in spList. Trading volume and daily price highs are stored for entries since the beginning of 2020. 
```{r}
volumes = list()
highs = list()
# for (i in 1:nrow(spList)) {
for (i in 1:500) {
  sym = as.character(spList[i,1])
  temp = av_get(symbol = sym, 'TIME_SERIES_DAILY', outputsize='full')
  volume_temp = list(temp$volume[which(temp$timestamp > '2020-01-01')])
  high_temp = list(temp$high[which(temp$timestamp > '2020-01-01')])
  volumes = rbind(volumes, volume_temp)
  highs = rbind(highs, high_temp)
  Sys.sleep(12)
}
```
Order time series data into columns with each variable representing a day. 
```{r}
sp_daily_volumes = data.frame(spList)
sp_daily_highs = data.frame(spList)
times = temp$timestamp[which(temp$timestamp > '2020-01-01')] 
for (i in 1:length(times)) {
  volume_temp = c()
  for (j in 1:500) {
    volume_temp = c(volume_temp, volumes[[j]][i])
  }
  sp_daily_volumes = cbind(sp_daily_volumes, volume_temp)
  high_temp = c()
  for (j in 1:500) {
    high_temp = c(high_temp, highs[[j]][i])
  }
  sp_daily_highs = cbind(sp_daily_highs, high_temp)
}
```
Rename columns according to dates. 
```{r}
col_names = c('Symbol', 'Name', 'Sector', as.character(times))
names(sp_daily_volumes) = col_names
names(sp_daily_highs) = col_names
```
Aggregate volume data by sector and calculate daily averages. 
```{r}
volume_avgs = data.frame(date = times)
ind = sp_daily_volumes %>%
  filter(Sector == 'Industrials')
health = sp_daily_volumes %>%
  filter(Sector == 'Health Care')
info = sp_daily_volumes %>%
  filter(Sector == 'Information Technology')
comm = sp_daily_volumes %>%
  filter(Sector == 'Communication Services')
con = sp_daily_volumes %>%
  filter(Sector == 'Consumer Discretionary')
ut = sp_daily_volumes %>%
  filter(Sector == 'Utilities')
fin = sp_daily_volumes %>%
  filter(Sector == 'Financials')
mat = sp_daily_volumes %>%
  filter(Sector == 'Materials')
rlest = sp_daily_volumes %>%
  filter(Sector == 'Real Estate')
constpl = sp_daily_volumes %>%
  filter(Sector == 'Consumer Staples')
nrg = sp_daily_volumes %>%
  filter(Sector == 'Energy')
mean_ind = c()
mean_health = c()
mean_info = c()
mean_comm = c()
mean_con = c()
mean_ut = c()
mean_fin = c()
mean_mat = c()
mean_rlest = c()
mean_constpl = c()
mean_nrg = c()
for (i in 4:(3+length(times))) {
  mean_ind = c(mean_ind, as.numeric(sapply(ind[i], mean)))
  mean_health = c(mean_health, as.numeric(sapply(health[i], mean)))
  mean_info = c(mean_info, as.numeric(sapply(info[i], mean)))
  mean_comm = c(mean_comm, as.numeric(sapply(comm[i], mean)))
  mean_con = c(mean_con, as.numeric(sapply(con[i], mean)))
  mean_ut = c(mean_ut, as.numeric(sapply(ut[i], mean)))
  mean_fin = c(mean_fin, as.numeric(sapply(fin[i], mean)))
  mean_mat = c(mean_mat, as.numeric(sapply(mat[i], mean)))
  mean_rlest = c(mean_rlest, as.numeric(sapply(rlest[i], mean)))
  mean_constpl = c(mean_constpl, as.numeric(sapply(constpl[i], mean)))
  mean_nrg = c(mean_nrg, as.numeric(sapply(nrg[i], mean)))
}
volume_avgs$mean_ind = mean_ind
volume_avgs$mean_health = mean_health
volume_avgs$mean_info = mean_info
volume_avgs$mean_comm = mean_comm
volume_avgs$mean_con = mean_con
volume_avgs$mean_ut = mean_ut
volume_avgs$mean_fin = mean_fin
volume_avgs$mean_mat = mean_mat
volume_avgs$mean_rlest = mean_rlest
volume_avgs$mean_constpl = mean_constpl
volume_avgs$mean_nrg = mean_nrg
```
Plot daily average trading volume by sector. 
```{r}
ggplot(volume_avgs) +
  geom_line(aes(x = date, y = mean_ind), col = 'red') +
  geom_line(aes(x = date, y = mean_health), col = 'blue') +
  geom_line(aes(x = date, y = mean_info), col = 'green') +
  geom_line(aes(x = date, y = mean_comm), col = 'orange') +
  geom_line(aes(x = date, y = mean_con), col = 'purple') +
  geom_line(aes(x = date, y = mean_ut), col = 'yellow') +
  geom_line(aes(x = date, y = mean_fin), col = 'turquoise') +
  geom_line(aes(x = date, y = mean_mat), col = 'pink') +
  geom_line(aes(x = date, y = mean_rlest), col = 'brown') +
  geom_line(aes(x = date, y = mean_constpl), col = 'dark blue') +
  geom_line(aes(x = date, y = mean_nrg), col = 'black') +
  labs(x = 'Date', y = 'Mean Sector Volume', title = 'Average Trading Volumes by Sector', subtitle = 'January 2020-Present')
```
Aggregate price high data by sector and calculate daily averages. 
```{r}
high_avgs = data.frame(date = times)
ind = sp_daily_highs %>%
  filter(Sector == 'Industrials')
health = sp_daily_highs %>%
  filter(Sector == 'Health Care')
info = sp_daily_highs %>%
  filter(Sector == 'Information Technology')
comm = sp_daily_highs %>%
  filter(Sector == 'Communication Services')
con = sp_daily_highs %>%
  filter(Sector == 'Consumer Discretionary')
ut = sp_daily_highs %>%
  filter(Sector == 'Utilities')
fin = sp_daily_highs %>%
  filter(Sector == 'Financials')
mat = sp_daily_highs %>%
  filter(Sector == 'Materials')
rlest = sp_daily_highs %>%
  filter(Sector == 'Real Estate')
constpl = sp_daily_highs %>%
  filter(Sector == 'Consumer Staples')
nrg = sp_daily_highs %>%
  filter(Sector == 'Energy')
mean_ind = c()
mean_health = c()
mean_info = c()
mean_comm = c()
mean_con = c()
mean_ut = c()
mean_fin = c()
mean_mat = c()
mean_rlest = c()
mean_constpl = c()
mean_nrg = c()
for (i in 4:(3+length(times))) {
  mean_ind = c(mean_ind, as.numeric(sapply(ind[i], mean)))
  mean_health = c(mean_health, as.numeric(sapply(health[i], mean)))
  mean_info = c(mean_info, as.numeric(sapply(info[i], mean)))
  mean_comm = c(mean_comm, as.numeric(sapply(comm[i], mean)))
  mean_con = c(mean_con, as.numeric(sapply(con[i], mean)))
  mean_ut = c(mean_ut, as.numeric(sapply(ut[i], mean)))
  mean_fin = c(mean_fin, as.numeric(sapply(fin[i], mean)))
  mean_mat = c(mean_mat, as.numeric(sapply(mat[i], mean)))
  mean_rlest = c(mean_rlest, as.numeric(sapply(rlest[i], mean)))
  mean_constpl = c(mean_constpl, as.numeric(sapply(constpl[i], mean)))
  mean_nrg = c(mean_nrg, as.numeric(sapply(nrg[i], mean)))
}
high_avgs$mean_ind = mean_ind
high_avgs$mean_health = mean_health
high_avgs$mean_info = mean_info
high_avgs$mean_comm = mean_comm
high_avgs$mean_con = mean_con
high_avgs$mean_ut = mean_ut
high_avgs$mean_fin = mean_fin
high_avgs$mean_mat = mean_mat
high_avgs$mean_rlest = mean_rlest
high_avgs$mean_constpl = mean_constpl
high_avgs$mean_nrg = mean_nrg
```
Plot daily average price high by sector. 
```{r}
ggplot(high_avgs) +
  geom_line(aes(x = date, y = mean_ind), col = 'red') +
  geom_line(aes(x = date, y = mean_health), col = 'blue') +
  geom_line(aes(x = date, y = mean_info), col = 'green') +
  geom_line(aes(x = date, y = mean_comm), col = 'orange') +
  geom_line(aes(x = date, y = mean_con), col = 'purple') +
  geom_line(aes(x = date, y = mean_ut), col = 'yellow') +
  geom_line(aes(x = date, y = mean_fin), col = 'turquoise') +
  geom_line(aes(x = date, y = mean_mat), col = 'pink') +
  geom_line(aes(x = date, y = mean_rlest), col = 'brown') +
  geom_line(aes(x = date, y = mean_constpl), col = 'dark blue') +
  geom_line(aes(x = date, y = mean_nrg), col = 'black') +
  labs(x = 'Date', y = 'Mean Sector High', title = 'Average Daily Price Highs by Sector', subtitle = 'January 2020-Present')
```