---
title: "STAT 472 Project Workflow"
author: "Adam Kiehl"
date: "5/10/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(alphavantager) # API Connection
library(RColorBrewer) # Colorings 
library(strucchange) # Chow Test
library(forecast) # ARIMA Models
library(tseries) # Autocorrelation Function
library(MaskJointDensity) # anyNA() Function
library(corrplot) # corrplot() Function
library(ggpubr) # ggarrange() Function
library(forecast) 
```

# Set-Up and Data Wrangling

Import and style scraped data frame of daily S&P 500 prices. 
```{r}
sp_daily_highs = read.csv('./data/sp_daily_highs_lg.csv') %>%
  select(-X)
times = names(sp_daily_highs)[4:length(names(sp_daily_highs))]
times = gsub('X', '', times)
times = gsub('\\.', '-', times)
times = as.Date(times)
col_names = c('Symbol', 'Name', 'Sector', as.character(times))
names(sp_daily_highs) = col_names
```
Aggregate price high data by sector and calculate daily averages. mean_tot is calculated as daily average of entire S&P 500. mean_groupa, mean_groupb, and mean_groupc are also calculated and the justification for those groupings is given later on. 
```{r}
high_avgs = data.frame(date = times)
ind = sp_daily_highs %>%
  filter(Sector == 'Industrials')
health = sp_daily_highs %>%
  filter(Sector == 'Health Care')
info = sp_daily_highs %>%
  filter(Sector == 'Information Technology')
comm = sp_daily_highs %>%
  filter(Sector == 'Communication Services')
con = sp_daily_highs %>%
  filter(Sector == 'Consumer Discretionary')
ut = sp_daily_highs %>%
  filter(Sector == 'Utilities')
fin = sp_daily_highs %>%
  filter(Sector == 'Financials')
mat = sp_daily_highs %>%
  filter(Sector == 'Materials')
rlest = sp_daily_highs %>%
  filter(Sector == 'Real Estate')
constpl = sp_daily_highs %>%
  filter(Sector == 'Consumer Staples')
nrg = sp_daily_highs %>%
  filter(Sector == 'Energy')
groupa = rbind(comm, con, info)
groupb = rbind(fin, ind, mat)
groupc = rbind(rlest, ut)
mean_ind = c()
mean_health = c()
mean_info = c()
mean_comm = c()
mean_con = c()
mean_ut = c()
mean_fin = c()
mean_mat = c()
mean_rlest = c()
mean_constpl = c()
mean_nrg = c()
mean_tot = c()
mean_groupa = c()
mean_groupb = c()
mean_groupc = c()
for (i in 4:(3+length(times))) {
  mean_ind = c(mean_ind, as.numeric(sapply(ind[i], mean)))
  mean_health = c(mean_health, as.numeric(sapply(health[i], mean)))
  mean_info = c(mean_info, as.numeric(sapply(info[i], mean)))
  mean_comm = c(mean_comm, as.numeric(sapply(comm[i], mean)))
  mean_con = c(mean_con, as.numeric(sapply(con[i], mean)))
  mean_ut = c(mean_ut, as.numeric(sapply(ut[i], mean)))
  mean_fin = c(mean_fin, as.numeric(sapply(fin[i], mean)))
  mean_mat = c(mean_mat, as.numeric(sapply(mat[i], mean)))
  mean_rlest = c(mean_rlest, as.numeric(sapply(rlest[i], mean)))
  mean_constpl = c(mean_constpl, as.numeric(sapply(constpl[i], mean)))
  mean_nrg = c(mean_nrg, as.numeric(sapply(nrg[i], mean)))
  mean_tot = c(mean_tot, as.numeric(sapply(sp_daily_highs[i], mean)))
  mean_groupa = c(mean_groupa, as.numeric(sapply(groupa[i], mean)))
  mean_groupb = c(mean_groupb, as.numeric(sapply(groupb[i], mean)))
  mean_groupc = c(mean_groupc, as.numeric(sapply(groupc[i], mean)))
}
high_avgs$mean_ind = mean_ind
high_avgs$mean_health = mean_health
high_avgs$mean_info = mean_info
high_avgs$mean_comm = mean_comm
high_avgs$mean_con = mean_con
high_avgs$mean_ut = mean_ut
high_avgs$mean_fin = mean_fin
high_avgs$mean_mat = mean_mat
high_avgs$mean_rlest = mean_rlest
high_avgs$mean_constpl = mean_constpl
high_avgs$mean_nrg = mean_nrg
high_avgs$mean_tot = mean_tot
high_avgs$groupa = mean_groupa
high_avgs$groupb = mean_groupb
high_avgs$groupc = mean_groupc
```
Plot daily average price highs by sector. 
```{r}
ggplot(high_avgs) +
  geom_line(aes(x = date, y = mean_ind, color = 'Industrials')) +
  geom_line(aes(x = date, y = mean_info, color = 'Information Technology')) +
  geom_line(aes(x = date, y = mean_comm, color = 'Communication Services')) +
  geom_line(aes(x = date, y = mean_con, color = 'Consumer Discretionary')) +
  geom_line(aes(x = date, y = mean_ut, color = 'Utilities')) +
  geom_line(aes(x = date, y = mean_fin, color = 'Financials')) +
  geom_line(aes(x = date, y = mean_mat, color = 'Materials')) +
  geom_line(aes(x = date, y = mean_rlest, color = 'Real Estate')) +
  geom_line(aes(x = date, y = mean_constpl, color = 'Consumer Staples')) +
  geom_line(aes(x = date, y = mean_nrg, color = 'Energy')) +
  geom_line(aes(x = date, y = mean_health, color = 'Health Care')) + 
  geom_line(aes(x = date, y = mean_tot, color = 'Total')) + 
  geom_vline(xintercept = as.Date('2020-03-11'), linetype = 3, alpha = .75) +
  geom_text(aes(x = as.Date('2020-03-11'), y = 315, label = 'WHO Declares COVID-19 Pandemic'), angle = 45, size = 2) + 
  scale_color_manual(values = c(
    'Communication Services' = brewer.pal(12, 'Set3')[1],
    'Consumer Discretionary' = brewer.pal(12, 'Set3')[2],
    'Consumer Staples' = brewer.pal(12, 'Set3')[3],
    'Energy' = brewer.pal(12, 'Set3')[4],
    'Financials' = brewer.pal(12, 'Set3')[5],
    'Health Care' = brewer.pal(12, 'Set3')[6],
    'Industrials' = brewer.pal(12, 'Set3')[7],
    'Information Technology' = brewer.pal(12, 'Set3')[8],
    'Materials' = brewer.pal(12, 'Set3')[10],
    'Real Estate' = brewer.pal(12, 'Set3')[11],
    'Total'= 'black',
    'Utilities' = brewer.pal(12, 'Set3')[12]
  )) +
  labs(x='Date', y='Mean Sector Price High', title='Average Price Highs by Sector', subtitle='January 2016-Present', color='Sector')

ggplot(high_avgs %>% filter(date > '2020-01-01')) +
  geom_line(aes(x = date, y = mean_ind, color = 'Industrials')) +
  geom_line(aes(x = date, y = mean_info, color = 'Information Technology')) +
  geom_line(aes(x = date, y = mean_comm, color = 'Communication Services')) +
  geom_line(aes(x = date, y = mean_con, color = 'Consumer Discretionary')) +
  geom_line(aes(x = date, y = mean_ut, color = 'Utilities')) +
  geom_line(aes(x = date, y = mean_fin, color = 'Financials')) +
  geom_line(aes(x = date, y = mean_mat, color = 'Materials')) +
  geom_line(aes(x = date, y = mean_rlest, color = 'Real Estate')) +
  geom_line(aes(x = date, y = mean_constpl, color = 'Consumer Staples')) +
  geom_line(aes(x = date, y = mean_nrg, color = 'Energy')) +
  geom_line(aes(x = date, y = mean_health, color = 'Health Care')) + 
  geom_line(aes(x = date, y = mean_tot, color = 'Total')) + 
  geom_vline(xintercept = as.Date('2020-03-11'), linetype = 3, alpha = .75) +
  geom_text(aes(x = as.Date('2020-03-11'), y = 315, label = 'WHO Declares COVID-19 Pandemic'), angle = 45, size = 2) + 
  scale_color_manual(values = c(
    'Communication Services' = brewer.pal(12, 'Set3')[1],
    'Consumer Discretionary' = brewer.pal(12, 'Set3')[2],
    'Consumer Staples' = brewer.pal(12, 'Set3')[3],
    'Energy' = brewer.pal(12, 'Set3')[4],
    'Financials' = brewer.pal(12, 'Set3')[5],
    'Health Care' = brewer.pal(12, 'Set3')[6],
    'Industrials' = brewer.pal(12, 'Set3')[7],
    'Information Technology' = brewer.pal(12, 'Set3')[8],
    'Materials' = brewer.pal(12, 'Set3')[10],
    'Real Estate' = brewer.pal(12, 'Set3')[11],
    'Total'= 'black',
    'Utilities' = brewer.pal(12, 'Set3')[12]
  )) +
  labs(x='Date', y='Mean Sector Price High', title='Average Price Highs by Sector', subtitle='January 2020-Present', color='Sector')
```
Write data frame to .csv for safe-keeping.
```{r}
write.csv(high_avgs, './data/high_avgs.csv')
```
Separate data into training and testing sets. Data before 2021 will be used for training. 
```{r}
high_train = high_avgs %>%
  filter(as.Date(date) < as.Date('2021-01-01'))
high_test = high_avgs %>%
  filter(as.Date(date) >= as.Date('2021-01-01'))
```

# Structural Breaks 

Identify strucutral breaks within 2020 data using Chow Test in strucchange package. 
```{r}
high_2020 = high_train %>%
  filter(as.Date(date) > '2020-01-01')
print(high_2020$date[breakpoints(high_2020$mean_tot ~ high_2020$date)$breakpoints])
```
Plot the time series to visualize structural breakpoints. The first break aligns closely with the WHO declaring the COVID-19 pandemic and the initial March shutdown. The second break corresponds to the week where many of the initial stay-at-home orders expired. The third break corresponds roughly with the start of the school and the onset of fall. 
```{r}
plt <- ggplot(high_2020) +
  geom_line(aes(x = date, y = mean_tot)) +
  geom_vline(xintercept = as.Date('2020-01-20'), linetype = 3, alpha = .75) +
  geom_text(aes(x = as.Date('2020-01-20'), y = 200, label = 'First US Case'), angle = 45, size = 2) + 
  geom_vline(xintercept = as.Date('2020-03-11'), linetype = 3, alpha = .75) +
  geom_text(aes(x = as.Date('2020-03-11'), y = 200, label = 'WHO Declares Pandemic'), angle = 45, size = 2) + 
  geom_vline(xintercept = as.Date('2020-03-26'), linetype = 3, alpha = .75) +
  geom_text(aes(x = as.Date('2020-03-26'), y = 200, label = 'CARES Act'), angle = 45, size = 2) + 
  geom_vline(xintercept = as.Date('2020-05-28'), linetype = 3, alpha = .75) +
  geom_text(aes(x = as.Date('2020-05-28'), y = 200, label = '100,000 US Deaths'), angle = 45, size = 2) + 
  geom_vline(xintercept = as.Date('2020-11-03'), linetype = 3, alpha = .75) +
  geom_text(aes(x = as.Date('2020-11-03'), y = 200, label = 'US Election'), angle = 45, size = 2) + 
  geom_vline(xintercept = as.Date('2020-11-18'), linetype = 3, alpha = .75) +
  geom_text(aes(x = as.Date('2020-11-18'), y = 200, label = '250,000 US Deaths'), angle = 45, size = 2) + 
  geom_vline(xintercept = as.Date('2020-12-11'), linetype = 3, alpha = .75) +
  geom_text(aes(x = as.Date('2020-12-11'), y = 200, label = 'Pfizer Vaccine'), angle = 45, size = 2) + 
  geom_vline(xintercept = as.Date('2020-12-18'), linetype = 3, alpha = .75) +
  geom_text(aes(x = as.Date('2020-12-18'), y = 200, label = 'Moderna Vaccine'), angle = 45, size = 2) + 
  geom_vline(xintercept = as.Date('2020-09-01'), linetype = 3, alpha = .75) +
  geom_text(aes(x = as.Date('2020-09-01'), y = 200, label = 'US Rejects WHO Vaccine Effort'), angle = 45, size = 2) + 
  labs(x='Date', y='Mean Price High', title='S&P 500 Average Price Highs', subtitle='January 2020-January 2021') +
  ylim(100, 250)
for (brk in breakpoints(high_2020$mean_tot ~ high_2020$date)$breakpoints) {
  plt <- plt + geom_vline(xintercept = as.Date(high_2020$date[brk]), linetype = 3, alpha = .75, color = 'red')
}

plt
```
Characterize state periods as defined by structural breaks. 
```{r, warning=FALSE}
prds_high = c(rep(1, 47), rep(2, 63), rep(3, 61), rep(4, 82))
high_2020 %>% 
  mutate(period = prds_high) %>%
  group_by(period) %>%
  summarize(Mean = mean(mean_tot), Var = var(mean_tot), .groups = 'keep')
```
Generate dummy variables and interactions for modelling break periods and seasons. 
```{r}
high_train$period1 = with(rle(high_train$date < '2020-03-11'), rep(as.integer(values), lengths))
high_train$period2 = with(rle((high_train$date >= '2020-03-11') & (high_train$date < '2020-06-10')), rep(as.integer(values), lengths))
high_train$period3 = with(rle((high_train$date >= '2020-06-10') & (high_train$date < '2020-09-04')), rep(as.integer(values), lengths))
high_train$period4 = with(rle(high_train$date >= '2020-09-04'), rep(as.integer(values), lengths))

high_train$interConstpl2 = high_train$mean_constpl * high_train$period2
high_train$interConstpl3 = high_train$mean_constpl * high_train$period3
high_train$interComm3 = high_train$mean_comm * high_train$period3

spring = c()
summer = c()
fall = c()
winter = c()
for (i in 1:length(high_train$date)) {
  mon = as.numeric(format(high_train$date[i], '%m'))
  if (mon %in% c(3, 4, 5)) {
    spring = c(spring, 1)
    summer = c(summer, 0)
    fall = c(fall, 0)
    winter = c(winter, 0)
  }
  else if (mon %in% c(6, 7, 8)) {
    spring = c(spring, 0)
    summer = c(summer, 1)
    fall = c(fall, 0)
    winter = c(winter, 0)
  }
  else if (mon %in% c(9, 10, 11)) {
    spring = c(spring, 0)
    summer = c(summer, 0)
    fall = c(fall, 1)
    winter = c(winter, 0)
  }
  else {
    spring = c(spring, 0)
    summer = c(summer, 0)
    fall = c(fall, 0)
    winter = c(winter, 1)
  }
}
```

# Pre-Modelling 

Create time series from sector averages with 252 trading days in a year.
```{r}
high_ts = data.frame(date = as.Date(high_train$date))
high_ts$high_ind_ts = ts(high_train$mean_ind, frequency = 252)
high_ts$high_health_ts = ts(high_train$mean_health, frequency = 252)
high_ts$high_info_ts = ts(high_train$mean_info, frequency = 252)
high_ts$high_comm_ts = ts(high_train$mean_comm, frequency = 252)
high_ts$high_con_ts = ts(high_train$mean_con, frequency = 252)
high_ts$high_ut_ts = ts(high_train$mean_ut, frequency = 252)
high_ts$high_fin_ts = ts(high_train$mean_fin, frequency = 252)
high_ts$high_mat_ts = ts(high_train$mean_mat, frequency = 252)
high_ts$high_rlest_ts = ts(high_train$mean_rlest, frequency = 252)
high_ts$high_constpl_ts = ts(high_train$mean_constpl, frequency = 252)
high_ts$high_nrg_ts = ts(high_train$mean_nrg, frequency = 252)
high_ts$high_tot_ts = ts(high_train$mean_tot, frequency = 252)
high_ts$high_groupa_ts = ts(high_train$groupa, frequency = 252)
high_ts$high_groupb_ts = ts(high_train$groupb, frequency = 252)
high_ts$high_groupc_ts = ts(high_train$groupc, frequency = 252)
high_ts$period2 = ts(high_train$period2, frequency = 252)
high_ts$period3 = ts(high_train$period3, frequency = 252)
high_ts$period4 = ts(high_train$period4, frequency = 252)
high_ts$interConstpl2 = ts(high_train$interConstpl2, frequency = 252)
high_ts$interConstpl3 = ts(high_train$interConstpl3, frequency = 252)
high_ts$interComm3 = ts(high_train$interComm3, frequency = 252)
high_ts$spring = ts(spring, frequency = 252)
high_ts$summer = ts(summer, frequency = 252)
high_ts$fall = ts(fall, frequency = 252)
```
These time series contain both seasonal and linear trends that must be removed to achieve stationarity. 
```{r}
plot(decompose(high_ts$high_tot_ts), xlab = 'Time [Years]')
```
Apply seasonal differencing transform across all time series to remove seasonality. An example plot is shown.
```{r}
high_tr1 = data.frame(date = as.Date(high_train$date[253:1259]))
high_tr1$high_ind_tr = diff(high_ts$high_ind_ts, lag = 252, differences = 1)
high_tr1$high_health_tr = diff(high_ts$high_health_ts, lag = 252, differences = 1)
high_tr1$high_info_tr = diff(high_ts$high_info_ts, lag = 252, differences = 1)
high_tr1$high_comm_tr = diff(high_ts$high_comm_ts, lag = 252, differences = 1)
high_tr1$high_con_tr = diff(high_ts$high_con_ts, lag = 252, differences = 1)
high_tr1$high_ut_tr = diff(high_ts$high_ut_ts, lag = 252, differences = 1)
high_tr1$high_fin_tr = diff(high_ts$high_fin_ts, lag = 252, differences = 1)
high_tr1$high_mat_tr = diff(high_ts$high_mat_ts, lag = 252, differences = 1)
high_tr1$high_rlest_tr = diff(high_ts$high_rlest_ts, lag = 252, differences = 1)
high_tr1$high_constpl_tr = diff(high_ts$high_constpl_ts, lag = 252, differences = 1)
high_tr1$high_nrg_tr = diff(high_ts$high_nrg_ts, lag = 252, differences = 1)
high_tr1$high_tot_tr = diff(high_ts$high_tot_ts, lag = 252, differences = 1)
high_tr1$high_groupa_tr = diff(high_ts$high_groupa_ts, lag = 252, differences = 1)
high_tr1$high_groupb_tr = diff(high_ts$high_groupb_ts, lag = 252, differences = 1)
high_tr1$high_groupc_tr = diff(high_ts$high_groupc_ts, lag = 252, differences = 1)
high_tr1$period2 = diff(high_ts$period2, lag = 252, differences = 1)
high_tr1$period3 = diff(high_ts$period3, lag = 252, differences = 1)
high_tr1$period4 = diff(high_ts$period4, lag = 252, differences = 1)
high_tr1$interConstpl2 = diff(high_ts$interConstpl2, lag = 252, differences = 1)
high_tr1$interConstpl3 = diff(high_ts$interConstpl3, lag = 252, differences = 1)
high_tr1$interComm3 = diff(high_ts$interComm3, lag = 252, differences = 1)
plot(high_tr1$high_tot_tr, main='Seasonally Transformed Time Series', xlab='Time [Years]', ylab='Price [Dollars]')
```
Apply first-differencing transform across all time series to remove trends. An example plot is shown. 
```{r}
high_tr2 = data.frame(date = as.Date(high_train$date[254:1259]))
high_tr2$high_ind_tr = diff(high_tr1$high_ind_tr, lag = 1, differences = 1)
high_tr2$high_health_tr = diff(high_tr1$high_health_tr, lag = 1, differences = 1)
high_tr2$high_info_tr = diff(high_tr1$high_info_tr, lag = 1, differences = 1)
high_tr2$high_comm_tr = diff(high_tr1$high_comm_tr, lag = 1, differences = 1)
high_tr2$high_con_tr = diff(high_tr1$high_con_tr, lag = 1, differences = 1)
high_tr2$high_ut_tr = diff(high_tr1$high_ut_tr, lag = 1, differences = 1)
high_tr2$high_fin_tr = diff(high_tr1$high_fin_tr, lag = 1, differences = 1)
high_tr2$high_mat_tr = diff(high_tr1$high_mat_tr, lag = 1, differences = 1)
high_tr2$high_rlest_tr = diff(high_tr1$high_rlest_tr, lag = 1, differences = 1)
high_tr2$high_constpl_tr = diff(high_tr1$high_constpl_tr, lag = 1, differences = 1)
high_tr2$high_nrg_tr = diff(high_tr1$high_nrg_tr, lag = 1, differences = 1)
high_tr2$high_tot_tr = diff(high_tr1$high_tot_tr, lag = 1, differences = 1)
high_tr2$high_groupa_tr = diff(high_tr1$high_groupa_tr, lag = 1, differences = 1)
high_tr2$high_groupb_tr = diff(high_tr1$high_groupb_tr, lag = 1, differences = 1)
high_tr2$high_groupc_tr = diff(high_tr1$high_groupc_tr, lag = 1, differences = 1)
high_tr2$period2 = diff(high_tr1$period2, lag = 1, differences = 1)
high_tr2$period3 = diff(high_tr1$period3, lag = 1, differences = 1)
high_tr2$period4 = diff(high_tr1$period4, lag = 1, differences = 1)
high_tr2$interConstpl2 = diff(high_tr1$interConstpl2, lag = 1, differences = 1)
high_tr2$interConstpl3 = diff(high_tr1$interConstpl3, lag = 1, differences = 1)
high_tr2$interComm3 = diff(high_tr1$interComm3, lag = 1, differences = 1)
plot(high_tr2$high_tot_tr, main='Seasonally, First Transformed Time Series', xlab='Time [Years]', ylab='Price [Dollars]')
```
Generate a correlation matrix between sectors using differenced time series. From this matrix, the following groups can be estimated: groupA- comm, con, info; groupB- fin, ind, mat; groupC- rlest, ut. 
```{r}
cor(high_tr2 %>% 
      select(-c(date, high_tot_tr, high_groupa_tr, high_groupb_tr, high_groupc_tr, period2, period3, period4, interConstpl2, interConstpl3, interComm3)) %>% 
      sapply(unlist), 
    high_tr2 %>% 
      select(-c(date, high_tot_tr, high_groupa_tr, high_groupb_tr, high_groupc_tr, period2, period3, period4, interConstpl2, interConstpl3, interComm3)) %>% 
      sapply(unlist))
```
Visualize correlations between sectors. 
```{r}
corrplot(cor(high_tr2 %>% 
      select(-c(date, high_tot_tr, high_groupa_tr, high_groupb_tr, high_groupc_tr, period2, period3, period4, interConstpl2, interConstpl3, interComm3)) %>% 
      sapply(unlist), 
    high_tr2 %>% 
      select(-c(date, high_tot_tr, high_groupa_tr, high_groupb_tr, high_groupc_tr, period2, period3, period4, interConstpl2, interConstpl3, interComm3)) %>% 
      sapply(unlist)))
```
Confirm groupings with exploratory factor analysis. Factor 1 is roughly equivalent to groupA. Factor 3 is roughly equivalent to groupB. Factor 2 is roughly equivalent to groupC. Factor 4 represents the energy sector, Factor 5 represents the consumer staples sector, and the health sector is here included with Factor 1. 
```{r}
factanal(high_tr2[,2:12,], 6)
```
Test time series for stability with Auto-Correlation Function plots using the Box-Jenkins method. 
```{r}
acf(high_tr2$high_groupa_tr, lag = 252, main='Group A ACF Plot')
acf(high_tr2$high_groupb_tr, lag = 252, main='Group B ACF Plot')
acf(high_tr2$high_groupc_tr, lag = 252, main='Group C ACF Plot')
acf(high_tr2$high_health_tr, lag = 252, main='Health Care ACF Plot')
acf(high_tr2$high_constpl_tr, lag = 252, main='Consumer Staples ACF Plot')
acf(high_tr2$high_nrg_tr, lag = 252, main='Energy ACF Plot')
```

# Modelling

Optimize AIC over ARMA paramaters p and q to tune order() with 0 differences built in. 
```{r}
final.aic = Inf
final.order = c(0, 0, 0)
for (p in 0:10) {
  for (q in 0:10) {
    tryCatch(
      {
        current.aic = AIC(Arima(high_tr2$high_tot_tr, order = c(p, 0, q), xreg = as.matrix(high_ts$period2[254:1259])))
        if (current.aic < final.aic) {
          final.aic = current.aic
          final.order = c(p, 0, q)
        }
      }, 
      error = function(cond) {
      }
    )
  }
}
final.order
```
Fit ARMA models for each sector grouping with dummy variable for the break period. 
```{r}
order = final.order
modelA1 = Arima(high_tr2$high_groupa_tr, 
               order = order, 
               xreg = as.matrix(high_ts$period2[254:1259]))
modelB1 = Arima(high_tr2$high_groupb_tr, 
               order = order, 
               xreg = as.matrix(high_ts$period2[254:1259]))
modelC1 = Arima(high_tr2$high_groupc_tr, 
               order = order, 
               xreg = as.matrix(high_ts$period2[254:1259]))
modelHealth1 = Arima(high_tr2$high_health_tr, 
                    order = order, 
                    xreg = as.matrix(high_ts$period2[254:1259]))
modelConstpl1 = Arima(high_tr2$high_constpl_tr,
                    order = order, 
                    xreg = as.matrix(high_ts$period2[254:1259]))
modelNrg1 = Arima(high_tr2$high_nrg_tr, 
                    order = order, 
                    xreg = as.matrix(high_ts$period2[254:1259]))
```
Fit ARMA models for each sector grouping with dummy variables for the break period and seasons. 
```{r}
order = final.order
modelA2 = Arima(diff(high_ts$high_groupa_ts, lag = 1, differences = 1), 
               order = order, 
               xreg = as.matrix(cbind(high_ts$period2[-1], high_ts$spring[-1], high_ts$summer[-1], high_ts$fall[-1])))
modelB2 = Arima(diff(high_ts$high_groupb_ts, lag = 1, differences = 1), 
               order = order, 
               xreg = as.matrix(cbind(high_ts$period2[-1], high_ts$spring[-1], high_ts$summer[-1], high_ts$fall[-1])))
modelC2 = Arima(diff(high_ts$high_groupc_ts, lag = 1, differences = 1), 
               order = order, 
               xreg = as.matrix(cbind(high_ts$period2[-1], high_ts$spring[-1], high_ts$summer[-1], high_ts$fall[-1])))
modelHealth2 = Arima(diff(high_ts$high_health_ts, lag = 1, differences = 1), 
                    order = order, 
                    xreg = as.matrix(cbind(high_ts$period2[-1], high_ts$spring[-1], high_ts$summer[-1], high_ts$fall[-1])))
modelConstpl2 = Arima(diff(high_ts$high_constpl_ts, lag = 1, differences = 1),
                    order = order, 
                    xreg = as.matrix(cbind(high_ts$period2[-1], high_ts$spring[-1], high_ts$summer[-1], high_ts$fall[-1])))
modelNrg2 = Arima(diff(high_ts$high_nrg_ts, lag = 1, differences = 1), 
                    order = order, 
                    xreg = as.matrix(cbind(high_ts$period2[-1], high_ts$spring[-1], high_ts$summer[-1], high_ts$fall[-1])))
```
Again, optimize AIC over ARMA paramaters p and q to tune order() with two differences built in. 
```{r}
final.aic = Inf
final.order = c(0, 0, 0)
for (p in 0:10) {
  for (q in 0:10) {
    tryCatch(
      {
        current.aic = AIC(Arima(high_ts$high_tot_ts, order = c(p, 2, q), xreg = high_ts$period2))
        if (current.aic < final.aic) {
          final.aic = current.aic
          final.order = c(p, 2, q)
        }
      }, 
      error = function(cond) {
      }
    )
  }
}

final.order
```
Fit ARMA model for total time series using second-order differencing model.  
```{r}
order = final.order
modelTot3 = Arima(high_ts$high_tot_ts, order = order, xreg = high_ts$period2)
```

# Forecasting

Forecast values according to seasonally and first transformed ARMA model. Take the more extreme of the upper and lower end of the forecast's confidence interval to better simulate the 'fizziness' of the market. An example forecast plot is shown. 
```{r}
futureXReg = as.matrix(rep(0, 60))

modelA_forecast = forecast(modelA1, xreg = as.matrix(futureXReg))[c('upper', 'lower')]
modelB_forecast = forecast(modelB1, xreg = as.matrix(futureXReg))[c('upper', 'lower')]
modelC_forecast = forecast(modelC1, xreg = as.matrix(futureXReg))[c('upper', 'lower')]
modelHealth_forecast = forecast(modelHealth1, xreg = as.matrix(futureXReg))[c('upper', 'lower')]
modelConstpl_forecast = forecast(modelConstpl1, xreg = as.matrix(futureXReg))[c('upper', 'lower')]
modelNrg_forecast = forecast(modelNrg1, xreg = as.matrix(futureXReg))[c('upper', 'lower')]

modelA1_pred = c()
modelB1_pred = c()
modelC1_pred = c()
modelHealth1_pred = c()
modelConstpl1_pred = c()
modelNrg1_pred = c()

for (i in 1:length(modelA_forecast$upper[,1])) {
  if (modelA_forecast$upper[i,2] > abs(modelA_forecast$lower[i,2])) {
    modelA1_pred = c(modelA1_pred, modelA_forecast$upper[i,2])
  }
  else {
    modelA1_pred = c(modelA1_pred, modelA_forecast$lower[i,2])
  }
  
  if (modelB_forecast$upper[i,2] > abs(modelB_forecast$lower[i,2])) {
    modelB1_pred = c(modelB1_pred, modelB_forecast$upper[i,2])
  }
  else {
    modelB1_pred = c(modelB1_pred, modelB_forecast$lower[i,2])
  }
  
  if (modelC_forecast$upper[i,2] > abs(modelC_forecast$lower[i,2])) {
    modelC1_pred = c(modelC1_pred, modelC_forecast$upper[i,2])
  }
  else {
    modelC1_pred = c(modelC1_pred, modelC_forecast$lower[i,2])
  }
  
  if (modelHealth_forecast$upper[i,2] > abs(modelHealth_forecast$lower[i,2])) {
    modelHealth1_pred = c(modelHealth1_pred, modelHealth_forecast$upper[i,2])
  }
  else {
    modelHealth1_pred = c(modelHealth1_pred, modelHealth_forecast$lower[i,2])
  }
  
  if (modelConstpl_forecast$upper[i,2] > abs(modelConstpl_forecast$lower[i,2])) {
    modelConstpl1_pred = c(modelConstpl1_pred, modelConstpl_forecast$upper[i,2])
  }
  else {
    modelConstpl1_pred = c(modelConstpl1_pred, modelConstpl_forecast$lower[i,2])
  }
  
  if (modelNrg_forecast$upper[i,2] > abs(modelNrg_forecast$lower[i,2])) {
    modelNrg1_pred = c(modelNrg1_pred, modelNrg_forecast$upper[i,2])
  }
  else {
    modelNrg1_pred = c(modelNrg1_pred, modelNrg_forecast$lower[i,2])
  }
}

plot(forecast(modelA1, xreg = as.matrix(futureXReg)), main='Seasonally, First Transformed ARMA(7, 0, 10) Errors', xlab='Time [Years]', ylab='Price [Dollars]')
```
Forecast values according to first-differenced and seasonally modelled ARMA model. Take the more extreme of the upper and lower end of the forecast's confidence interval to better simulate the 'fizziness' of the market. An example forecast plot is shown. 
```{r}
futureXReg = as.matrix(cbind(rep(0, 60), c(rep(0, 42), rep(1, 18)), rep(0, 60), rep(0, 60)))

modelA_forecast = forecast(modelA2, xreg = as.matrix(futureXReg))[c('upper', 'lower')]
modelB_forecast = forecast(modelB2, xreg = as.matrix(futureXReg))[c('upper', 'lower')]
modelC_forecast = forecast(modelC2, xreg = as.matrix(futureXReg))[c('upper', 'lower')]
modelHealth_forecast = forecast(modelHealth2, xreg = as.matrix(futureXReg))[c('upper', 'lower')]
modelConstpl_forecast = forecast(modelConstpl2, xreg = as.matrix(futureXReg))[c('upper', 'lower')]
modelNrg_forecast = forecast(modelNrg2, xreg = as.matrix(futureXReg))[c('upper', 'lower')]

modelA2_pred = c()
modelB2_pred = c()
modelC2_pred = c()
modelHealth2_pred = c()
modelConstpl2_pred = c()
modelNrg2_pred = c()

for (i in 1:length(modelA_forecast$upper[,1])) {
  if (modelA_forecast$upper[i,2] > abs(modelA_forecast$lower[i,2])) {
    modelA2_pred = c(modelA2_pred, modelA_forecast$upper[i,2])
  }
  else {
    modelA2_pred = c(modelA2_pred, modelA_forecast$lower[i,2])
  }
  
  if (modelB_forecast$upper[i,2] > abs(modelB_forecast$lower[i,2])) {
    modelB2_pred = c(modelB2_pred, modelB_forecast$upper[i,2])
  }
  else {
    modelB2_pred = c(modelB2_pred, modelB_forecast$lower[i,2])
  }
  
  if (modelC_forecast$upper[i,2] > abs(modelC_forecast$lower[i,2])) {
    modelC2_pred = c(modelC2_pred, modelC_forecast$upper[i,2])
  }
  else {
    modelC2_pred = c(modelC2_pred, modelC_forecast$lower[i,2])
  }
  
  if (modelHealth_forecast$upper[i,2] > abs(modelHealth_forecast$lower[i,2])) {
    modelHealth2_pred = c(modelHealth2_pred, modelHealth_forecast$upper[i,2])
  }
  else {
    modelHealth2_pred = c(modelHealth2_pred, modelHealth_forecast$lower[i,2])
  }
  
  if (modelConstpl_forecast$upper[i,2] > abs(modelConstpl_forecast$lower[i,2])) {
    modelConstpl2_pred = c(modelConstpl2_pred, modelConstpl_forecast$upper[i,2])
  }
  else {
    modelConstpl2_pred = c(modelConstpl2_pred, modelConstpl_forecast$lower[i,2])
  }
  
  if (modelNrg_forecast$upper[i,2] > abs(modelNrg_forecast$lower[i,2])) {
    modelNrg2_pred = c(modelNrg2_pred, modelNrg_forecast$upper[i,2])
  }
  else {
    modelNrg2_pred = c(modelNrg2_pred, modelNrg_forecast$lower[i,2])
  }
}

plot(forecast(modelA2, xreg = as.matrix(futureXReg)), main='First Transformed, Seasonally Modelled ARMA(7, 0, 10) Errors', xlab='Time [Years]', ylab='Price [Dollars]')
```
Forecast and plot values according to twice-differenced ARMA model.
```{r}
futureXReg = as.matrix(rep(0, 60))

modelTot3_pred = forecast(modelTot3, xreg = as.matrix(futureXReg))$mean

plot(forecast(modelTot3, xreg = as.matrix(futureXReg)), main='Twice-Differenced ARMA(9, 2, 3) Errors', xlab='Time [Years]', ylab='Price [Dollars]')
```
Use Ljung-Box test to validate models lagged by twice the supposed seasonal period. 
```{r}
lag = 504

print('Seasonally and First Differenced:')
Box.test(resid(modelA1), lag = lag, type = 'Ljung-Box')$p.value
Box.test(resid(modelB1), lag = lag, type = 'Ljung-Box')$p.value
Box.test(resid(modelC1), lag = lag, type = 'Ljung-Box')$p.value
Box.test(resid(modelHealth1), lag = lag, type = 'Ljung-Box')$p.value
Box.test(resid(modelConstpl1), lag = lag, type = 'Ljung-Box')$p.value
Box.test(resid(modelNrg1), lag = lag, type = 'Ljung-Box')$p.value

print('First Differenced and Seasonally Modelled:')
Box.test(resid(modelA2), lag = lag, type = 'Ljung-Box')$p.value
Box.test(resid(modelB2), lag = lag, type = 'Ljung-Box')$p.value
Box.test(resid(modelC2), lag = lag, type = 'Ljung-Box')$p.value
Box.test(resid(modelHealth2), lag = lag, type = 'Ljung-Box')$p.value
Box.test(resid(modelConstpl2), lag = lag, type = 'Ljung-Box')$p.value
Box.test(resid(modelNrg2), lag = lag, type = 'Ljung-Box')$p.value

print('Twice-Differenced:')
Box.test(resid(modelTot3), lag = lag, type = 'Ljung-Box')$p.value
```
Use cumulative sum method to forecast re-adjsted future values from last known training point. A fourth series is created by averaging other three forecasts.
```{r}
y = high_ts$high_tot_ts
pred = rowMeans(cbind(modelA1_pred, modelB1_pred, modelC1_pred, modelHealth1_pred, modelConstpl1_pred, modelNrg1_pred))
for (t in 1260:1319) {
  trans = pred[t-1259] + y[t-252] - y[t-253] + y[t-1]
  y = c(y, trans)
}
cumsumSeries1 = y[1260:1319]

cumsumSeries2 = cumsum(c(high_ts$high_tot_ts[1259], 
                        rowMeans(
                              cbind(
                                modelA2_pred, modelB2_pred, modelC2_pred, modelHealth2_pred, modelConstpl2_pred, modelNrg2_pred
                                ))))[-1]

cumsumSeries3 = modelTot3_pred

cumsumSeries4 = rowMeans(cbind(cumsumSeries1, cumsumSeries2, cumsumSeries3))
```
Calculate MSE for each prediction set.
```{r}
mean((high_avgs$mean_tot[which(as.Date(high_avgs$date) >= '2021-01-01')][1:60] - cumsumSeries1)^2)
mean((high_avgs$mean_tot[which(as.Date(high_avgs$date) >= '2021-01-01')][1:60] - cumsumSeries2)^2)
mean((high_avgs$mean_tot[which(as.Date(high_avgs$date) >= '2021-01-01')][1:60] - cumsumSeries3)^2)
mean((high_avgs$mean_tot[which(as.Date(high_avgs$date) >= '2021-01-01')][1:60] - cumsumSeries4)^2)
```
Plot aggregated time series and mean_tot time series that it simulates. 
```{r}
pred1plot = ggplot(mapping = aes(x = as.Date(high_avgs$date)[which(as.Date(high_avgs$date) > '2019-01-01')])) +
  geom_path(aes(y = high_avgs$mean_tot[which(as.Date(high_avgs$date) > '2019-01-01')])) +
  geom_path(aes(y = c(high_ts$high_tot_ts[which(as.Date(high_ts$date) > '2019-01-01')], cumsumSeries1)), color = 'red') +
  labs(title = 'Seasonally and First-Differenced', subtitle = 'MSE: 862.04', x = 'Date', y = 'Price [$]')
pred2plot = ggplot(mapping = aes(x = as.Date(high_avgs$date)[which(as.Date(high_avgs$date) > '2019-01-01')])) +
  geom_path(aes(y = high_avgs$mean_tot[which(as.Date(high_avgs$date) > '2019-01-01')])) +
  geom_path(aes(y = c(high_ts$high_tot_ts[which(as.Date(high_ts$date) > '2019-01-01')], cumsumSeries2)), color = 'red') +
  labs(title = 'First-Differenced, Seasonally Modeled', subtitle = 'MSE: 392.37', x = 'Date', y = 'Price [$]')
pred3plot = ggplot(mapping = aes(x = as.Date(high_avgs$date)[which(as.Date(high_avgs$date) > '2019-01-01')])) +
  geom_path(aes(y = high_avgs$mean_tot[which(as.Date(high_avgs$date) > '2019-01-01')])) +
  geom_path(aes(y = c(high_ts$high_tot_ts[which(as.Date(high_ts$date) > '2019-01-01')], cumsumSeries3)), color = 'red') +
  labs(title = 'Twice-Differenced Predictions', subtitle = 'MSE: 32.04', x = 'Date', y = 'Price [$]')
pred4plot = ggplot(mapping = aes(x = as.Date(high_avgs$date)[which(as.Date(high_avgs$date) > '2019-01-01')])) +
  geom_path(aes(y = high_avgs$mean_tot[which(as.Date(high_avgs$date) > '2019-01-01')])) +
  geom_path(aes(y = c(high_ts$high_tot_ts[which(as.Date(high_ts$date) > '2019-01-01')], cumsumSeries4)), color = 'red') +
  labs(title = 'Forecast Combination Predictions', subtitle = 'MSE: 301.66', x = 'Date', y = 'Price [$]')
ggarrange(pred1plot, pred2plot, pred3plot, pred4plot)
```



